Supplementary Material\\ \textit{Cross Validation Approaches for Penalized Cox Regression}

### 2019-05-23 main.tex
Very minor edits to 3.4
change to basic; minor edits to 3.2

\note{I hate to change the name again, but what do you think about calling it ``basic'' cross-validation instead of ``standard''?  My concern is that it's not really standard in the sense of ``the thing most people do''; I think most people use V\&VH since that's the default in glmnet.}

@@

\par One approach is to simply calculate the partial likelihood based on the observations in the test set as a measure for the model's predictive accuracy.  In doing so, we calculate the risk for each member of the test set relative to other members of the test set.  One drawback to this approach is that it becomes unstable when the size of the the test set is small.  In particular, it cannot be applied to leave-one-out cross-validation (LOOCV), as we must have at least two observations in the test set to compare their risk relative to each other.

@@

\par A direct approach to measure a model's predictive accuracy in $D_k$ is to evaluate the log partial likelihood at $\hat{\beta}^{-k}$ using the data in $D_k$. Cross-validation error (CVE) is then measured by the total deviance across all $K$ folds, which (up to a constant) is given by:
\begin{equation}
  \label{eq:standard}
  \CVE = -2 \sum_{k=1}^{K} \ell^{k}(\hat{\beta}^{-k});
\end{equation}
the factor of 2 is arbitrary, but often used so that CVE estimates the deviance of the model.
We refer this approach as \emph{standard cross-validated partial likelihood}. This is implemented in the $\tt{glmnet}$ package as the \verb|"ungrouped"| option.

\par The standard approach is by far the most common way of conducting cross-validation for models such as linear regression and logistic regression. However, the standard approach can be problematic for Cox regression in the sense that that there may not be enough observations in $D_k$ to build up the risk set for partial likelihood. For example, the standard approach cannot work with leave-one-out cross-validation, since for all folds, $\ell_k$ would either be zero or undefined.

To address this issue, \citet{Verweij1993} proposed an alternative method for cross-validation in Cox models:
\begin{equation}
\label{eq:VVH}
	\CVE = -2\sum_{k = 1}^K \left\{ \ell(\hat{\beta}^{- k})  - \ell^{-k}(\hat{\beta}^{- k}) \right\}. 
\end{equation}
Here, $\ell(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ using the entire data set $D$ and $\ell^{-k}(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ on
the training set $T_k$. By avoiding the construction of a partial likelihood on the test set, this approach ensures that the risk set is always sufficiently large and the partial likelihood well-defined.  We refer to this cross-validation method as the Verweij and Van Houwelingen (V\&VH) approach. For many models, such as logistic regression or linear regression, the quantities $\ell(\hat{\beta}^{- k})  - \ell^{-k}(\hat{\beta}^{- k})$ and  $\ell^{k}(\hat{\beta}^{-k})$ are equivalent to each other -- in other words, the standard and V\&VH approaches agree.  However, Verweij and Van Houwelingen's approach is more stable for Cox regression as there is always a large number of observations in the risk sets its calculations are based on.  Since its proposal, the V\&VH approach has been widely implemented as a tool for cross-validation is Cox models.  For example, it is used by the R packages {\tt CoxRidge} \citep{CoxRidge}, {\tt fastcox} \citep{fastcox}, {\tt SGL} \citep{SGL} , {\tt CoxBoost} \citep{CoxBoost}, {\tt mboost} \citep{mboost}, and {\tt glmnet} \citep{glmnet}.  In {\tt glmnet}, it is the default choice for penalized Cox regression and is referred to as the \verb|"grouped"| option in the package's syntax.

@@

The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model. Hence, another approach to cross-validation would be to include an extra step of estimating the baseline hazard and using it to quantify the cross-validation error. The standard approach to estimating the cumulative baseline hazard, $\Lambda_0$, is the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
\begin{equation}
  \hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}.
\end{equation}
The Cox model implies the following relationship between the baseline cumulative hazard and the cumulative hazard for individual $i$:
\begin{equation}
  \hat{\Lambda}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\Tr\hat\beta).
\end{equation}

@@

In this section, we consider the accuracy of the cross-validation approaches introduced in Section~\ref{Sec:methods}. We varied the sample size, dimension, censoring percentage \note{(we don't really vary censoring percentage, do we?)}, and signal strength of the simulated data sets to examine how these factors affect the performance of cross-validation. Accuracy was assesed via several criteria.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[height= 7cm ]{./manuscript_figure/figure_2_new.png}
  \caption{\label{Fig:mse-brier-c} Simulation comparing 4 CV methods. The horizontal axis in all three plots is $s$, which represents the signal strength in the data. The value of $s$ is varied from 0.4 to 0.9.  $\log(\text{MSE ratio})$, relative to the oracle modle is plotted in the left panel. Out-of-sample Brier scores are plotted in the middle. Out-of-sample C index is plotted in the right panel. For each simulated data set, n = 120, p = 1000. Expected censoring percentage is 10$\%$. \note{For the sake of consistency, Signal should be capitalized in the x-axis label above.}}
\end{figure}

For our first simulation study, the covariate matrix has $n = 120$ observations and $p = 1000$ covariates. The covariate vector $\beta$ is assumed to be sparse, with $\beta_j = 0$ for $j \in \{11, 12, \ldots, 1000\}$ and $\beta_j = s$ for $j \in \{1, 2, \ldots, 10 \}$, where $s$ is a prespecified constant that specifies the signal strength in the simulated data set. Here, we vary $s$ from 0.4 to 0.9. The survival outcomes have 10$\%$ expected censoring. For each of the four methods described in Section~\ref{Sec:methods}, 10-fold cross-validation was implemented. $N = 200$ replications were used for each scenario.  Figure~\ref{Fig:mse-brier-c} illustrates the results of this simulation study. For all four metrics, the overall pattern is the same (KL scores not shown in Figure~\ref{Fig:mse-brier-c} due to space constraints, but they were very similar to Brier scores).  Although the four cross-validation methods performed relatively similarly, the standard and linear predictor approaches had the best performance across all metrics: lower MSE, better Brier scores, and higher C Index.

\note{Just curious: I would imagine that standard/basic cross-validation does best when there is least censoring, as in Section 3.2.  Is this true?  In other words, does Figure 2 make basic cross-validation look better than it often is, because the censoring rate is so small, just 10\%?}

\begin{table}[htb]

@@

\caption{\label{Tab:sim} Simulation comparing four cross-validation methods with varying dimension and signal strength}
\centering
\begin{tabular}[t]{cccccccccc}
\toprule
 n& p& $\#$ Non-Zero & Signal & Method & $\lambda$ & log(MSE Ratio) &Brier Score & Kullback Leibler & C Index\\
\midrule
150 & 50 & 5 & weak & V VH  & 0.098 (0.030) & 1.561 (0.851) & 0.245 (0.014) & 0.685 (0.030) & 0.610 (0.035) \\
    &    &   &     & Standard  & 0.088 (0.025) & 1.531 (0.838) & 0.244 (0.013) & 0.683 (0.029) & 0.614 (0.029) \\
    &    &   &     & LinearPred  & 0.092 (0.029) & 1.543 (0.846) & 0.245 (0.014) & 0.684 (0.030) & 0.612 (0.034) \\
    &    &   &     & DevResid  & 0.098 (0.029) & 1.547 (0.848) & 0.245 (0.014) & 0.684 (0.030) & 0.611 (0.035) \\
\addlinespace
150 & 50 & 5 & strong & V VH  & 0.071 (0.010) & 1.586 (0.989) & 0.176 (0.011) & 0.526 (0.027) & 0.750 (0.012) \\
    &    &   &     & Standard  & 0.066 (0.011) & 1.544 (0.971) & 0.176 (0.011) & 0.526 (0.028) & 0.749 (0.013) \\
    &    &   &     & LinearPred  & 0.067 (0.010) & 1.540 (0.981) & 0.176 (0.011) & 0.525 (0.028) & 0.750 (0.012) \\
    &    &   &     & DevResid  & 0.094 (0.011) & 1.995 (0.841) & 0.181 (0.011) & 0.539 (0.026) & 0.751 (0.013) \\
\addlinespace
400 & 10000 & 20 & weak & V VH  & 0.126 (0.019) & 2.958 (0.416) & 0.251 (0.012) & 0.696 (0.026) & 0.589 (0.044) \\
    &    &   &     & Standard  & 0.103 (0.022) & 2.880 (0.437) & 0.244 (0.015) & 0.680 (0.031) & 0.603 (0.039) \\
    &    &   &     & LinearPred  & 0.105 (0.023) & 2.886 (0.437) & 0.244 (0.015) & 0.681 (0.031) & 0.602 (0.039) \\
    &    &   &     & DevResid  & 0.107 (0.022) & 2.895 (0.433) & 0.245 (0.014) & 0.683 (0.03) & 0.602 (0.039) \\
\addlinespace
400 & 10000 & 20 & strong & V VH  & 0.084 (0.012) & 3.799 (0.548) & 0.176 (0.03) & 0.528 (0.069) & 0.763 (0.048) \\
    &    &   &     & Standard  & 0.059 (0.007) & 3.455 (0.578) & 0.154 (0.019) & 0.468 (0.049) & 0.780 (0.031) \\
    &    &   &     & LinearPred  & 0.060 (0.007) & 3.479 (0.574) & 0.155 (0.019) & 0.470 (0.049) & 0.780 (0.030) \\
    &    &   &     & DevResid  & 0.076 (0.018) & 3.708 (0.518) & 0.166 (0.024) & 0.502 (0.057) & 0.772 (0.049) \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{Tab:sim} displays the results of additional simulations carried out as we varied the dimension of the data set along with the signal strength.  Specifically, we examined the following settings: low dimension (p = 50) with weak signal, low dimension with strong signal, high dimension (p = 10000) with weak signal, and high dimension with strong signal. In all settings, the censoring rate was 30\%.  \note{What is s for the strong and weak cases?}

As in Figure~\ref{Fig:mse-brier-c}, the standard and linear predictor approaches outperform the deviance residual and V\&VH approaches.  The improvement is particularly noticeable in the most difficult setting: high-dimensional data with weak signal.  Table~\ref{Tab:sim} also provides some insight into why the deviance residual and V\&VH approaches perform poorly relative to standard CV and the linear predictor approach: they act conservatively, consistently choosing larger $\lam$ values and thus, selecting fewer variables.

@@

First, the number of uncensored observations in the test set may be insufficient to construct a well-defined partial likelihood for standard cross-validation, leaving the overall estimate of cross-validation error undefined as well. When computing the partial likelihood, only individuals whose events are observed will contribute to the likelihood. If there are no events observed in a particular fold, then the partial likelihood for that fold is undefined.

To illustrate how common this problem is, we generated 1000 independent data sets of $n = 100$ observations with a fixed censoring rate. For each data set, we conducted 10-fold cross-validation and recorded whether or not there was at least one event in each fold (i.e., whether the partial likelihood was well-defined for all folds). We repeated the same mechanism for scenarios with different censoring rates, ranging from $40\%$ to $80\%$. As illustrated in the left panel in Figure~\ref{Fig:stability}, at $n=100$ the standard approach can be problematic even at moderate censoring rates. For example, standard CV cannot be used with over 20\% of data sets when $70\%$ of observations are censored.

\begin{figure}[h]
  \centering
  \includegraphics[height= 7cm ]{./manuscript_figure/figure_3.png}
  \caption{\label{Fig:stability} Simulations illustrating the stability issues with the standard approach. For the figure on the left, $n = 100$, fold assignments are unbalanced, and the vertical axis is the proportion of undefined partial likelihoods. For the figure on the right, $n = 120$, fold assignments are balanced so that the number of events in each fold is the same, and the vertical axis is the standard deviation of the squared error loss. Basic cross-validation is much more variability than other approaches when there are only few events in each fold.}
\end{figure}

One approach for alleviating this issue is to balance the censoring status across all folds when making fold assignments. That is, to ensure the number of events observed is the same (or within 1) across all folds. Unless the number of the folds exceeds the number of observed events, such as in leave-one-out cross-validation, this approach will guarantee the partial likelihood in each fold is well-defined for the standard approach, thus the overall cross-validated error is well-defined.

In this case, however, even though the partial likelihood is well-defined, basic cross-validation suffers from extreme variability.  To illustrate, we generated data sets with $n = 120$ observations and a fixed censoring rate of $50\%$, so that there are 60 observed events. For each data set, we carried out cross-validation using balanced fold assignments with the number of folds set to be 10, 12, 15, 20, 30 and 60. When the number of folds was set to be 10, there were exactly 6 events in each fold; when the number of folds was set to be 60, there was exactly one event in each fold, and so on. After each fold assignment, we computed the cross-validated error using all four approaches described in Section~\ref{Sec:accuracy}. We recorded the model that minimizes the cross-validation error for each approach and computed its squared error loss. We replicated the whole process for $N = 200$ time. In the end, we computed the standard deviation of the squared error loss for each approach. \note{Does this phenomenon show up in the MSE too?}

The results of this simulation are illustrated in the right panel of Figure~\ref{Fig:stability}. While the other three approaches are not affected by the number of folds and number of events in each fold, the variability of basic cross-validation increases dramatically as the number of events per fold decreases. When there are sufficient number of events in each fold, the standard approach is similar to the other three approaches. However, when there are only one or two events in each fold, the standard approach can be up to 10 times more variable than other methods.  Thus, although the approach can be used to calculate CVE, there is considerable uncertainty surrounding its choice of the best model.

@@

\note{Should we have this paragraph?  Seems to contradict our conclusion, in which our recommendation is to use the linear predictor approach.}  Based on the two simulations illustrated in this section, we would make the following recommendations: i) always balance the censoring status when conducting fold assignments for the standard approach, ii) avoid using the standard approach in conjuction with high number of folds, iii) avoid using the standard approach when there are few number of events observed.

@@

As discussed earlier, standard cross-validation cannot be used to carry out leave-one-out cross validation (LOOCV) for Cox regression models. The other three approaches (V\&VH, linear predictor, and deviance residual), however, are not subject to the stability issues described in Section~\ref{Sec:stability}, and can be used to implement LOOCV.  Here, we compare these three approaches for carrying out LOOCV, and also compare LOOCV to 10-fold CV.

\begin{figure}[h]
  \centering
  \includegraphics[height= 8cm ]{./manuscript_figure/figure_4.png}
  \caption{\label{Fig:loocv}Simulation comparing LOO and 10-fold CV across three different approaches. With $N = 200$ replications, the data sets generated have dimension $n = 120$ and $p = 1000$. The vertical axis is the $\log$(MSE ratio). In each panel, the LOO version and 10-fold version of the same approach are plotted. In all three panels, 10-fold standard cross-validation is included as a baseline for comparison (blue line). \note{Again, Signal should be capitalized.}}
\end{figure}

For this simulation experiment, we generated datasets with $n = 120$ and $p = 1000$, using the same data generating procedure as described in Section~\ref{Sec:accuracy}. For each of the V\&VH, linear predictor, and deviance residual approaches, we conducted both 10-fold CV and LOOCV. We also conducted the standard 10-fold cross-validation as a baseline for comparison. We recorded the model that minimizes the cross-validated error  for each approach. After $N = 200$ replications, $\log$ of the MSE ratio was computed using the same definition described in Section~\ref{Sec:accuracy}.

Results of the simulation are shown in Figure~\ref{Fig:loocv}. The panels illustrate the results for each CV approach, one with the 10-fold version and one with the leave-one-out version. The 10-fold standard approach was plotted in all three panels as a baseline for comparison.  For all three approaches, the Leave-One-Out version selected models with a smaller $\log(\text{MSE ratio})$ than the 10-fold version. For the V\&VH and deviance residual approaches, however, LOOCV was still inferior to standard 10-fold CV.  As described in Section~\ref{Sec:accuracy}, these approaches tend to be conservative, selecting a larger-than-optimal value of the regularization parameter; LOOCV helps with this problem, but does not alleviate it entirely.

On the other hand, the LOOCV version of the linear predictor approach surpasses standard 10-fold CV, with a smaller $\log(\text{MSE ratio})$ across all signal strength levels. At the same time, even the 10-fold version of the linear predictor approach outperforms the leave-one-out version of the V$\&$VH approach and the deviance residual approach. In other words, although adopting a leave-one-out approach improves the predictive accuracy for V$\&$VH and deviance residual cross-validation, the magnitude of improvement in predictive performance is smaller than the improvement gained from switching to the linear predictor approach.

@@

\begin{figure}[h]
  \centering
  \includegraphics[height= 15cm ]{./manuscript_figure/figure_auc.png}
  \caption{\label{Fig:cvauc}Comparing Partial Likelihood and C Index both built from Cross-validated Linear Predictors. The horizontal axis represents the strength of signal $s$ in the data. The left figures are under the low dimension scenarios where p = 50 and n = 150. The right figures are under the high dimension scenarios where p = 1000 and n = 120. The expected percentages of censoring for both scenarios are 30 $\%$. The top row are box plots of the selected tuning parameter$\lambda$. The mid row plots the $\log(\text{relative Squared Error Loss})$, with respect to the oracle model. The bottom row plots the Brier Scores. }
\end{figure}

@@

\par We conducted simulations in both low dimensional ($p = 50$, $n = 150$) and high dimensional ($p = 1000$, $n = 120$) settings. For each of the scenario, we generated data from a spectrum of signal strength as in Section~\ref{Sec:accuracy}. Results of the simulations are illustrated in Figure~\ref{Fig:cvauc}.  We used the linear predictor approach (CV-LP) here as the representative approach for likelihood-based cross-validation.

@@

Cross-validation is a crutial step in fitting penalized Cox regression models. In this paper, we propose two cross-validation methods for Cox Regression: cross-validated linear predictors and cross-validated deviance residuals. We conducted a series of simulation experiments to study and compare their performance with two existing approaches, the standard cross-validation and Verweij and Van Houwelingen's cross-validated partial likelihood, in selecting tuning parameters when fitting penalized regression models.

>>>>

@@

\par One basic approach is to simply calculate the partial likelihood based on the observations in the test set as a measure for the model's predictive accuracy.  In doing so, we calculate the risk for each member of the test set relative to other members of the test set.  One drawback to this approach is that it becomes unstable when the size of the the test set is small.  In particular, it cannot be applied to leave-one-out cross-validation (LOOCV), as we must have at least two observations in the test set to compare their risk relative to each other.

@@

\par A direct approach to measure a model's predictive accuracy in $D_k$ is to evaluate the log partial likelihood at $\hat{\beta}^{-k}$ using the data in $D_k$. Cross-validation error (CVE) is then measured by the total deviance across all $K$ folds, which (up to a constant) is given by:
\begin{equation}
  \label{eq:basic}
  \CVE = -2 \sum_{k=1}^{K} \ell^{k}(\hat{\beta}^{-k});
\end{equation}
the factor of 2 is arbitrary, but often used so that CVE estimates the deviance of the model.
We refer this approach as \emph{basic cross-validated partial likelihood}. This is implemented in the $\tt{glmnet}$ package as the \verb|"ungrouped"| option.

\par The basic approach is by far the most common way of conducting cross-validation for models such as linear regression and logistic regression. However, the basic approach can be problematic for Cox regression in the sense that that there may not be enough observations in $D_k$ to build up the risk set for partial likelihood. For example, the basic approach cannot work with leave-one-out cross-validation, since for all folds, $\ell_k$ would either be zero or undefined.

To address this issue, \citet{Verweij1993} proposed an alternative method for cross-validation in Cox models:
\begin{equation}
\label{eq:VVH}
	\CVE = -2\sum_{k = 1}^K \left\{ \ell(\hat{\beta}^{- k})  - \ell^{-k}(\hat{\beta}^{- k}) \right\}. 
\end{equation}
Here, $\ell(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ using the entire data set $D$ and $\ell^{-k}(\hat{\beta}^{-k})$ is the log partial likelihood evaluated at $\hat{\beta}^{-k}$ on
the training set $T_k$. By avoiding the construction of a partial likelihood on the test set, this approach ensures that the risk set is always sufficiently large and the partial likelihood well-defined.  We refer to this cross-validation method as the Verweij and Van Houwelingen (V\&VH) approach. For many models, such as logistic regression or linear regression, the quantities $\ell(\hat{\beta}^{- k})  - \ell^{-k}(\hat{\beta}^{- k})$ and  $\ell^{k}(\hat{\beta}^{-k})$ are equivalent to each other -- in other words, the basic and V\&VH approaches agree.  However, Verweij and Van Houwelingen's approach is more stable for Cox regression as there is always a large number of observations in the risk sets its calculations are based on.  Since its proposal, the V\&VH approach has been widely implemented as a tool for cross-validation is Cox models.  For example, it is used by the R packages {\tt CoxRidge} \citep{CoxRidge}, {\tt fastcox} \citep{fastcox}, {\tt SGL} \citep{SGL} , {\tt CoxBoost} \citep{CoxBoost}, {\tt mboost} \citep{mboost}, and {\tt glmnet} \citep{glmnet}.  In {\tt glmnet}, it is the default choice for penalized Cox regression and is referred to as the \verb|"grouped"| option in the package's syntax.

@@

The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model. Hence, another approach to cross-validation would be to include an extra step of estimating the baseline hazard and using it to quantify the cross-validation error. The basic approach to estimating the cumulative baseline hazard, $\Lambda_0$, is the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
\begin{equation}
  \hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}.
\end{equation}
The Cox model implies the following relationship between the baseline cumulative hazard and the cumulative hazard for individual $i$:
\begin{equation}
  \hat{\Lambda}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\Tr\hat\beta).
\end{equation}

@@

In this section, we consider the accuracy of the cross-validation approaches introduced in Section~\ref{Sec:methods}. We varied the sample size, dimension, censoring percentage, and signal strength of the simulated data sets to examine how these factors affect the performance of cross-validation. Accuracy was assesed via several criteria.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[height= 7cm ]{./manuscript_figure/figure_2_new.png}
  \caption{\label{Fig:mse-brier-c} Simulation comparing 4 CV methods. The horizontal axis in all three plots is $s$, which represents the signal strength in the data. The value of $s$ is varied from 0.4 to 0.9.  $\log(\text{MSE ratio})$, relative to the oracle modle is plotted in the left panel. Out-of-sample Brier scores are plotted in the middle. Out-of-sample C index is plotted in the right panel. For each simulated data set, n = 120, p = 1000. Expected censoring percentage is 10$\%$.}
\end{figure}

For our first simulation study, the covariate matrix has $n = 120$ observations and $p = 1000$ covariates. The covariate vector $\beta$ is assumed to be sparse, with $\beta_j = 0$ for $j \in \{11, 12, \ldots, 1000\}$ and $\beta_j = s$ for $j \in \{1, 2, \ldots, 10 \}$, where $s$ is a prespecified constant that specifies the signal strength in the simulated data set. Here, we vary $s$ from 0.4 to 0.9. The survival outcomes have 10$\%$ expected censoring. For each of the four methods described in Section~\ref{Sec:methods}, 10-fold cross-validation was implemented. $N = 200$ replications were used for each scenario.  Figure~\ref{Fig:mse-brier-c} illustrates the results of this simulation study. For all four metrics, the overall pattern is the same (KL scores not shown in Figure~\ref{Fig:mse-brier-c} due to space constraints, but they were very similar to Brier scores).  Although the four cross-validation methods performed relatively similarly, the basic and linear predictor approaches had the best performance across all metrics: lower MSE, better Brier scores, and higher C Index.

\begin{table}[!htb]

@@

\caption{\label{Tab:sim} Simulation comparing four cross-validation methods with varying dimension and signal strength}
\centering
\begin{tabular}[t]{cccccccccc}
\toprule
 n& p& $\#$ Non-Zero & Signal & Method & $\lambda$ & log(MSE Ratio) &Brier Score & Kullback Leibler & C Index\\
\midrule
150 & 50 & 5 & weak & V\&VH  & 0.098 (0.030) & 1.561 (0.851) & 0.245 (0.014) & 0.685 (0.030) & 0.610 (0.035) \\
    &    &   &     & Basic  & 0.088 (0.025) & 1.531 (0.838) & 0.244 (0.013) & 0.683 (0.029) & 0.614 (0.029) \\
    &    &   &     & LinearPred  & 0.092 (0.029) & 1.543 (0.846) & 0.245 (0.014) & 0.684 (0.030) & 0.612 (0.034) \\
    &    &   &     & DevResid  & 0.098 (0.029) & 1.547 (0.848) & 0.245 (0.014) & 0.684 (0.030) & 0.611 (0.035) \\
\addlinespace
150 & 50 & 5 & strong & V\&VH  & 0.071 (0.010) & 1.586 (0.989) & 0.176 (0.011) & 0.526 (0.027) & 0.750 (0.012) \\
    &    &   &     & Basic  & 0.066 (0.011) & 1.544 (0.971) & 0.176 (0.011) & 0.526 (0.028) & 0.749 (0.013) \\
    &    &   &     & LinearPred  & 0.067 (0.010) & 1.540 (0.981) & 0.176 (0.011) & 0.525 (0.028) & 0.750 (0.012) \\
    &    &   &     & DevResid  & 0.094 (0.011) & 1.995 (0.841) & 0.181 (0.011) & 0.539 (0.026) & 0.751 (0.013) \\
\addlinespace
400 & 10000 & 20 & weak & V\&VH  & 0.126 (0.019) & 2.958 (0.416) & 0.251 (0.012) & 0.696 (0.026) & 0.589 (0.044) \\
    &    &   &     & Basic  & 0.103 (0.022) & 2.880 (0.437) & 0.244 (0.015) & 0.680 (0.031) & 0.603 (0.039) \\
    &    &   &     & LinearPred  & 0.105 (0.023) & 2.886 (0.437) & 0.244 (0.015) & 0.681 (0.031) & 0.602 (0.039) \\
    &    &   &     & DevResid  & 0.107 (0.022) & 2.895 (0.433) & 0.245 (0.014) & 0.683 (0.03) & 0.602 (0.039) \\
\addlinespace
400 & 10000 & 20 & strong & V\&VH  & 0.084 (0.012) & 3.799 (0.548) & 0.176 (0.03) & 0.528 (0.069) & 0.763 (0.048) \\
    &    &   &     & Basic  & 0.059 (0.007) & 3.455 (0.578) & 0.154 (0.019) & 0.468 (0.049) & 0.780 (0.031) \\
    &    &   &     & LinearPred  & 0.060 (0.007) & 3.479 (0.574) & 0.155 (0.019) & 0.470 (0.049) & 0.780 (0.030) \\
    &    &   &     & DevResid  & 0.076 (0.018) & 3.708 (0.518) & 0.166 (0.024) & 0.502 (0.057) & 0.772 (0.049) \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{Tab:sim} displays the results of additional simulations carried out as we varied the dimension of the data set along with the signal strength.  Specifically, we examined the following settings: low dimension (p = 50) with weak signal, low dimension with strong signal, high dimension (p = 10000) with weak signal, and high dimension with strong signal. In all settings, the censoring rate was 30\%. For weak cases, signal strength $s$ was set to be 0.3; for strong cases, $s$ was set to be 0.6.

As in Figure~\ref{Fig:mse-brier-c}, the basic and linear predictor approaches outperform the deviance residual and V\&VH approaches.  The improvement is particularly noticeable in the most difficult setting: high-dimensional data with weak signal.  Table~\ref{Tab:sim} also provides some insight into why the deviance residual and V\&VH approaches perform poorly relative to basic CV and the linear predictor approach: they act conservatively, consistently choosing larger $\lam$ values and thus, selecting fewer variables.

@@

First, the number of uncensored observations in the test set may be insufficient to construct a well-defined partial likelihood for basic cross-validation, leaving the overall estimate of cross-validation error undefined as well. When computing the partial likelihood, only individuals whose events are observed will contribute to the likelihood. If there are no events observed in a particular fold, then the partial likelihood for that fold is undefined.

To illustrate how common this problem is, we generated 1000 independent data sets of $n = 100$ observations with a fixed censoring rate. For each data set, we conducted 10-fold cross-validation and recorded whether or not there was at least one event in each fold (i.e., whether the partial likelihood was well-defined for all folds). We repeated the same mechanism for scenarios with different censoring rates, ranging from $40\%$ to $80\%$. As illustrated in the left panel in Figure~\ref{Fig:stability}, at $n=100$ the basic approach can be problematic even at moderate censoring rates. For example, basic CV cannot be used with over 20\% of data sets when $70\%$ of observations are censored.

\begin{figure}[!htb]
  \centering
  \includegraphics[height= 7cm ]{./manuscript_figure/figure_3.png}
  \caption{\label{Fig:stability} Simulations illustrating the stability issues with the basic approach. For the figure on the left, $n = 100$, fold assignments are unbalanced, and the vertical axis is the proportion of undefined partial likelihoods. For the figure on the right, $n = 120$, fold assignments are balanced so that the number of events in each fold is the same, and the vertical axis is the basic deviation of the squared error loss. Basic cross-validation is much more variability than other approaches when there are only few events in each fold.}
\end{figure}

One approach for alleviating this issue is to balance the censoring status across all folds when making fold assignments. That is, to ensure the number of events observed is the same (or within 1) across all folds. Unless the number of the folds exceeds the number of observed events, such as in leave-one-out cross-validation, this approach will guarantee the partial likelihood in each fold is well-defined for the basic approach, thus the overall cross-validated error is well-defined.

In this case, however, even though the partial likelihood is well-defined, basic cross-validation suffers from extreme variability.  To illustrate, we generated data sets with $n = 120$ observations and a fixed censoring rate of $50\%$, so that there are 60 observed events. For each data set, we carried out cross-validation using balanced fold assignments with the number of folds set to be 10, 12, 15, 20, 30 and 60. When the number of folds was set to be 10, there were exactly 6 events in each fold; when the number of folds was set to be 60, there was exactly one event in each fold, and so on. After each fold assignment, we computed the cross-validated error using all four approaches described in Section~\ref{Sec:accuracy}. We recorded the model that minimizes the cross-validation error for each approach and computed its squared error loss. We replicated the whole process for $N = 200$ time. In the end, we computed the standard deviation of the squared error loss for each approach.

The results of this simulation are illustrated in the right panel of Figure~\ref{Fig:stability}. While the other three approaches are not affected by the number of folds and number of events in each fold, the variability of basic cross-validation increases dramatically as the number of events per fold decreases. When there are sufficient number of events in each fold, the basic approach is similar to the other three approaches. However, when there are only one or two events in each fold, the basic approach can be up to 10 times more variable than other methods. Thus, although the approach can be used to calculate CVE, there is considerable uncertainty surrounding its choice of the best model. The same phenomenon is also observed for the mean squared error. When there are only few events in each fold, MSE of the basic approach is affected and increases with the increasing uncertainty.

@@

Based on the simulation results illustrated in this section, we suggest that one should proceed with caution if they want to use the basic approach: i) always balance the censoring status when conducting fold assignments for the basic approach, ii) avoid using the basic approach in conjuction with high number of folds, iii) avoid using the basic approach when there are few number of events observed. Overall we would recommend using the linear predictor approach than the basic approach in practice.

@@

As discussed earlier, basic cross-validation cannot be used to carry out leave-one-out cross validation (LOOCV) for Cox regression models. The other three approaches (V\&VH, linear predictor, and deviance residual), however, are not subject to the stability issues described in Section~\ref{Sec:stability}, and can be used to implement LOOCV.  Here, we compare these three approaches for carrying out LOOCV, and also compare LOOCV to 10-fold CV.

\begin{figure}[!htb]
  \centering
  \includegraphics[height= 8cm ]{./manuscript_figure/figure_4.png}
  \caption{\label{Fig:loocv}Simulation comparing LOO and 10-fold CV across three different approaches. With $N = 200$ replications, the data sets generated have dimension $n = 120$ and $p = 1000$. The vertical axis is the $\log$(MSE ratio). In each panel, the LOO version and 10-fold version of the same approach are plotted. In all three panels, 10-fold basic cross-validation is included as a baseline for comparison (blue line).}
\end{figure}

For this simulation experiment, we generated datasets with $n = 120$ and $p = 1000$, using the same data generating procedure as described in Section~\ref{Sec:accuracy}. For each of the V\&VH, linear predictor, and deviance residual approaches, we conducted both 10-fold CV and LOOCV. We also conducted the basic 10-fold cross-validation as a baseline for comparison. We recorded the model that minimizes the cross-validated error  for each approach. After $N = 200$ replications, $\log$ of the MSE ratio was computed using the same definition described in Section~\ref{Sec:accuracy}.

Results of the simulation are shown in Figure~\ref{Fig:loocv}. The panels illustrate the results for each CV approach, one with the 10-fold version and one with the leave-one-out version. The 10-fold basic approach was plotted in all three panels as a baseline for comparison.  For all three approaches, the Leave-One-Out version selected models with a smaller $\log(\text{MSE ratio})$ than the 10-fold version. For the V\&VH and deviance residual approaches, however, LOOCV was still inferior to basic 10-fold CV.  As described in Section~\ref{Sec:accuracy}, these approaches tend to be conservative, selecting a larger-than-optimal value of the regularization parameter; LOOCV helps with this problem, but does not alleviate it entirely.

On the other hand, the LOOCV version of the linear predictor approach surpasses basic 10-fold CV, with a smaller $\log(\text{MSE ratio})$ across all signal strength levels. At the same time, even the 10-fold version of the linear predictor approach outperforms the leave-one-out version of the V$\&$VH approach and the deviance residual approach. In other words, although adopting a leave-one-out approach improves the predictive accuracy for V$\&$VH and deviance residual cross-validation, the magnitude of improvement in predictive performance is smaller than the improvement gained from switching to the linear predictor approach.

@@

@@

\par We conducted simulations in both low dimensional ($p = 50$, $n = 150$) and high dimensional ($p = 1000$, $n = 120$) settings. For each scenario, we generated data from a spectrum of signal strength as in Section~\ref{Sec:accuracy}. Results of the simulations are illustrated in Figure~\ref{Fig:cvauc}.  We used the linear predictor approach (CV-LP) here as the representative approach for likelihood-based cross-validation.

\begin{figure}[!htb]
  \centering
  \includegraphics[height= 15cm ]{./manuscript_figure/figure_auc.png}
  \caption{\label{Fig:cvauc}Comparing Partial Likelihood and C Index both built from Cross-validated Linear Predictors. The horizontal axis represents the strength of signal $s$ in the data. The left figures are under the low dimension scenarios where p = 50 and n = 150. The right figures are under the high dimension scenarios where p = 1000 and n = 120. The expected percentages of censoring for both scenarios are 30 $\%$. The top row are box plots of the selected tuning parameter$\lambda$. The mid row plots the $\log(\text{relative Squared Error Loss})$, with respect to the oracle model. The bottom row plots the Brier Scores. }
\end{figure}

@@

Cross-validation is a crutial step in fitting penalized Cox regression models. In this paper, we propose two cross-validation methods for Cox Regression: cross-validated linear predictors and cross-validated deviance residuals. We conducted a series of simulation experiments to study and compare their performance with two existing approaches, the basic cross-validation and Verweij and Van Houwelingen's cross-validated partial likelihood, in selecting tuning parameters when fitting penalized regression models.
